Modular flow and holographic entanglement entropy in cutoff AdS

### 2021-05-27 alpha/CutoffModularFlow/cutoffModular-alpha.tex
Update note on local entanglement entropy

@@

	\item We try to obtain the entanglement entropy using the modular flow at the \mbox{codim-2} cutoff surface, by defining a \textit{local entanglement entropy $\beta(x)$}, which relates the entanglement entropy to some thermal entropy.

@@

\pagebreak
\subsection{Modular flow for the cutoff field theory}
\textbf{Summary:} We try to obtain the entanglement entropy using the modular flow at the \mbox{codim-2} cutoff surface, by defining a \textit{local entanglement entropy $\beta(x)$}, which relates the entanglement entropy to some thermal entropy.

@@

	We note that there is a trick to read out the entanglement entropy from the modular flow: it is possible to define a \textit{local entanglement temperature}, and then use it to read the entanglement entropy from the modular flow generator, by comparing with the Hamiltonian of a thermal field theory, where we have:
	\begin{equation}
	\mcal{H}=\int \beta\,T_{00}\,dx,\quad \beta = 2\pi,
	\quad \xi|_{t=0} = \beta\,\pdd{t}
	\label{eq:beta_thermal}
	\end{equation}
	In general, the modular Hamiltonian is given by:
	\begin{equation}
	\mcal{H}_\mcal{A}=\int \beta(x)\, T_{00}\,dx,\quad \xi|_{t=0} = \beta(x)\, \pdd{t}
	\end{equation}

	Let's see this shortcut in action in CFT$_2$; the \textit{entropy density} is given by:
	\begin{equation}
	\partial_x S = \frac{c}{3} \frac{\pi}{\beta(x)}
	\end{equation}
	in terms of the local entanglement temperature $\beta(x)$, and the total entropy of the interval $\mcal{A}$ is given by:
	\begin{equation}
	S= \frac{c}{3}
		\int_\mcal{A} \frac{\pi}{\beta(x)}dx
	\end{equation}

	For example, for a single interval in CFT$_2$,
	\begin{equation}
	\beta(x)=\frac{2\pi}{L_\infty}
		\pqty{\frac{L_\infty^2}{4}-x^2}
	\end{equation}
	\begin{equation}
	S= \frac{c}{3} \int_{-L_\infty/2}^{L_\infty/2}
			\frac{L_\infty/2}{(L_\infty/2)^2-x^2}dx,
	\end{equation}
	The indefinite integral $
		\int \frac{L_\infty/2}{(L_\infty/2)^2-x^2}dx
		= \mop{arctanh}\frac{2x}{L_\infty}
		\rightarrow\infty
	$, as $x\rightarrow L_\infty/2$, so one needs a cutoff:
	\begin{equation}
	S=\frac{c}{3}\int_{-L_\infty/2+a}^{L_\infty/2-a}
		\frac{L_\infty/2}{(L_\infty/2)^2-x^2}dx
	=\frac{2c}{3} \mop{arctanh} \pqty{
			1-\frac{2a}{L_\infty}
		}
	\end{equation}
	As $a\rightarrow 0$,
	\begin{equation}
		\mop{arctanh} \pqty{
			1-\frac{2a}{L_\infty}
		}\rightarrow \frac{1}{2}\log\frac{L_\infty}{a}
	\end{equation}
	So we get the usual result
	\begin{equation}
	S=\frac{c}{3}\log\frac{L_\infty}{a}
	\end{equation}

	Another case is the Rindler patch,
	\begin{equation}
	\xi\propto x\partial_x-t\partial_x
	\end{equation}
	\begin{equation}
	\beta(x)=2\pi x
	\label{eq:beta_rindler}
	\end{equation}
	\begin{equation}
	S=\int_{a}^{l}\frac{c}{6x}=\frac{c}{6}\log \frac{l}{a}
	\end{equation}
	The factor $\frac{c}{6}$ is that in this case, one only considers one of the endpoints instead of two. One can read the C-function from \eqref{eq:beta_thermal}\eqref{eq:beta_rindler}, when $\beta$ is linear of $x$.
	\begin{equation}
	C=\frac{c}{6}
	\end{equation}

@@

	Now for the cutoff field theory living at the boundary, consider a modular flow generator of this form,
	\begin{equation}
	\xi|_{t=0}=\frac{2\pi}{L_\infty}(L_\infty^2/4-x^2)\partial_t
	\end{equation}
	Namely in the same form as the modular flow of a CFT, but we have:
	\begin{equation}
	L^2/4+a^2=L_\infty^2/4
	\end{equation}
	Where the interval for the cutoff theory is of length $L$. The entanglement entropy is
	\begin{equation}
	S=\frac{c}{6}\mop{arctanh}\frac{2x}{L_\infty}\Big|_{-L/2}^{L/2}
	=\frac{c}{3}\mop{arctanh}\frac{L}{\sqrt{L^2+4a^2}}
	=\frac{c}{3}\mop{arcsinh}\frac{L}{2a}
	\end{equation}
	with C-function
	\begin{equation}
	C(l)=\frac{c}{3}\frac{1}{\sqrt{1+\frac{4a^2}{L^2}}}
	\end{equation}
	Which agrees with \cite{Lewkowycz:2019xse}.

	In the field theory side, such modular flow generator means that the Rindler transformation map a slightly large region $D(\tilde{A})$ to the hyperbolic cylinder, and keep track of the endpoints of the interval $A$. The endpoints are implied in the range of the integration in the local entanglement temperature method, and also in the length of the cylinder in the Rindler map method.
	\begin{equation}
	M:D(A)\rightarrow R\times S
	\end{equation}
	\begin{equation}
	\tilde{M}:D(\tilde{A})\rightarrow R\times S
	\end{equation}
	A is the interval with length $L$, $\tilde{A}$ is the interval with length $L_\infty$.

@@

\textit{
	The discussion above is in the framework of $\mrm{CFT}_2$, but with different modular flow generator. Now we consider the framework of $T\bar{T}$ deformed theories. The thermal entropy with unit length is
	\begin{equation}
	S_{thermal}=\frac{\pi c}{3\beta}\eta
	\end{equation}
	\begin{equation}
	\eta=\frac{1}{\sqrt{1+\frac{4\pi^2c\mu}{3\beta^2}}}
	\end{equation}
	Consider the half line case,
	\begin{equation}
	\beta(x)=2\pi x
	\end{equation}
	One can get the C-function,
	\begin{equation}
	C(l)=\frac{c}{6}\frac{1}{\sqrt{1+\frac{c\mu}{3l^2}}}
	\end{equation}
	Again, there is ``one endpoint'' in this case, which will give one half contribution of the C-function with the pre-factor $c/3$.
}

>>>>

\newcommand{\lui}[2]{\textcolor{red}{#1}\todo[color=yellow]{\scriptsize{L: #2}}}

@@

	\item We try to obtain the entanglement entropy using the modular flow at the \mbox{codim-2} cutoff surface, using the \textit{effective inverse temperature $\beta(x)$} introduced in \cite{Wong:2013gua,Cardy:2016fqc}, which relates the entanglement entropy to some thermal entropy. \sidenote{[relation with partial entanglement entropy?]}

	\item \sidenote{We should look at the boosted interval.}

@@

\pagebreak
\subsection{Modular flow for the cutoff field theory}
\textbf{Summary:} We try to obtain the entanglement entropy using the modular flow at the \mbox{codim-2} cutoff surface, using the \textit{effective inverse temperature $\beta(x)$} introduced in \cite{Cardy:2016fqc}, which relates the entanglement entropy to some thermal entropy.

@@

	We note that there is a trick to read out the entanglement entropy from the modular flow: it is possible to define a \textit{local entanglement temperature}, and then use it to read the entanglement entropy from the modular flow generator \cite{Cardy:2016fqc}. The result is obtained by comparing with the Hamiltonian of a thermal field theory, where we have:
	\begin{equation}
	\mcal{H}=\int \beta\,T_{00}\,dx,\quad \beta = 2\pi,
	\quad \xi|_{t=0} = \beta\,\pdd{t}
	\label{eq:beta_thermal}
	\end{equation}
	In general, the modular Hamiltonian is given by:
	\begin{equation}
	\mcal{H}_\mcal{A}=\int \beta(x)\, T_{00}\,dx,\quad \xi|_{t=0} = \beta(x)\, \pdd{t}
	\end{equation}

	Let's see this trick in action in CFT$_2$; the \textit{entropy density} is given by:
	\begin{equation}
	s_\mcal{A} = \frac{c}{3} \frac{\pi}{\beta(x)}
	\end{equation}
	in terms of the local entanglement temperature $\beta(x)$, and the total entropy of the interval $\mcal{A}$ is given by:
	\begin{equation}
	S= \frac{c}{3}
		\int_\mcal{A} \frac{\pi}{\beta(x)}dx
	\end{equation}

	For example, for a single interval in CFT$_2$,
	\begin{equation}
	\beta(x)=\frac{2\pi}{L_\infty}
		\pqty{\frac{L_\infty^2}{4}-x^2}
	\end{equation}
	\begin{equation}
	S= \frac{c}{3} \int_{-L_\infty/2}^{L_\infty/2}
			\frac{L_\infty/2}{(L_\infty/2)^2-x^2}dx,
	\end{equation}
	The indefinite integral $
		\int \frac{L_\infty/2}{(L_\infty/2)^2-x^2}dx
		= \mop{arctanh}\frac{2x}{L_\infty}
%		\rightarrow\infty
	$ diverges as $x\rightarrow L_\infty/2$, so one needs a cutoff:
	\begin{equation}
	S=\frac{c}{3}\int_{-L_\infty/2+a}^{L_\infty/2-a}
		\frac{L_\infty/2}{(L_\infty/2)^2-x^2}dx
	=\frac{2c}{3} \mop{arctanh} \pqty{
			1-\frac{2a}{L_\infty}
		}
	\end{equation}
	As $a\rightarrow 0$,
	\begin{equation}
		\mop{arctanh} \pqty{
			1-\frac{2a}{L_\infty}
		}\rightarrow \frac{1}{2}\log\frac{L_\infty}{a}
	\end{equation}
	So we get the usual result
	\begin{equation}
	S=\frac{c}{3}\log\frac{L_\infty}{a}
	\end{equation}

	Another case is the Rindler patch, \sidenote{better to integrate from $l$ to $\infty$},
	\begin{equation}
	\xi\propto x\partial_x-t\partial_x
	\end{equation}
	\begin{equation}
	\beta(x)=2\pi x
	\label{eq:beta_rindler}
	\end{equation}
	\begin{equation}
	S=\int_{a}^{l}\frac{c}{6x}=\frac{c}{6}\log \frac{l}{a}
	\end{equation}
	The factor $\frac{c}{6}$ is that in this case, one only considers one of the endpoints instead of two. One can read the C-function from \eqref{eq:beta_thermal}\eqref{eq:beta_rindler}, when $\beta$ is linear of $x$.
	\begin{equation}
	C=\frac{c}{6}
	\end{equation}

@@

	Now for the cutoff field theory living at the boundary, consider a modular flow generator of this form,
	\begin{equation}
	\xi|_{t=0}=\frac{2\pi}{L_\infty}(L_\infty^2/4-x^2)\partial_t
	\end{equation}
	Namely in the same form as the modular flow of a CFT, but we have:
	\begin{equation}
	L^2/4+z_c^2=L_\infty^2/4
	\end{equation}
	Where the interval for the cutoff theory is of length $L$. The entanglement entropy is
	\begin{equation}
	S=\frac{c}{6}\mop{arctanh}\frac{2x}{L_\infty}\Big|_{x=-L/2}^{x=L/2}
	=\frac{c}{3}\mop{arctanh}\frac{L}{\sqrt{L^2+4z_c^2}}
	=\frac{c}{3}\mop{arcsinh}\frac{L}{2z_c}
	\end{equation}
	with C-function
	\begin{equation}
	C(l)=\frac{c}{3}\frac{1}{\sqrt{1+\frac{4z_c^2}{L^2}}}
	\end{equation}
	Which agrees with \cite{Lewkowycz:2019xse}. \sidenote{This provides a field theory side calculation of the cutoff entanglement entropy, based on the input of the modular flow $\xi$. }

	\textit{Note that we only need to know the modular flow $\xi$ along some time slice. We don't need to know the full modular flow. This seems to be consistent with Lewkowycz--Maldacena \cite{Lewkowycz:2013nqa} and the holographic entropy, where it only depends linearly on $\xi$ along the RT surface, as long as $\xi$ is a Killing vector. See \eqref{kdef}. }

	In the field theory side, such modular flow generator means that the Rindler transformation map a slightly large region $D(\tilde{A})$ to the hyperbolic cylinder, and keep track of the endpoints of the interval $A$. The endpoints are implied in the range of the integration in the local entanglement temperature method, and also in the length of the cylinder in the Rindler map method.
	\begin{equation}
	f:D(A)\rightarrow R\times S
	\end{equation}
	\begin{equation}
	\tilde{f}:D(\tilde{A})\rightarrow R\times S
	\end{equation}
	A is the interval with length $L$, $\tilde{A}$ is the interval with length $L_\infty$.

@@

\subsubsection*{Reasoning for this trick}
	The validity of this trick can be argued from the Rindler map; suppose we have the following Rindler map at the boundary:
	\begin{equation}
	\begin{aligned}
		f\colon
		\mcal{D}(\mcal{A})
		&\longto \mcal{R} = S^1_\beta \times f(\mcal{A}), \\
		(t,x)
		&\longmapsto (\tau,\tilde{x}),
	\end{aligned}
	\end{equation}
	Here $f(\mcal{A})$ is the image of the boundary interval. $f$ by construction ``normalizes'' the temperature for the modular generator, namely, the modular flow $\xi$ is now realized as the pullback of $2\pi\pdd{\tau}$:
	\begin{equation}
		\xi = \xi^\mu \pdd{\mu}
		= 2\pi\pdd{\tau}
	\end{equation}
	In particular, along the interval $\mcal{A}$ at the $t = 0$ time slice, we have:
	\begin{equation}
		\xi|_{t = 0}
		= \beta(x)\,\pdd{t}
		= 2\pi\pdd{\tau},
	\quad
		\pdv{t}{\tau}
		= \frac{\beta(x)}{2\pi}
	\end{equation}
	Note that although $\xi$ is a Killing vector when it's extended to the bulk, it is only a \textit{conformal Killing vector} when it's restricted to the boundary.

	Here we think of $f$ as a simple coordinates transformation or diffeomorphism, which keeps the metric $\dd{s}^2$ invariant, and maps the flat coordinates to a conformally flat coordinates:
	\begin{equation}
		\dd{s}^2
		= \dd{t}^2 + \dd{x}^2
		= \Omega^2(\tau,\tilde{x}) \pqty{
				\dd{\tau}^2 + \dd{\tilde{x}}^2
			},
	\quad
		\Omega_{t = 0} = \pdv{t}{\tau}
		= \frac{\beta(x)}{2\pi}
	\end{equation}

	Due to conformal symmetry, one can further drop the conformal factor $\Omega^2(\tau,\tilde{x})$ from the metric, while the thermal entropy of the system remains unchanged. We now have a flat annulus:
	\begin{equation}
		\tilde{\mcal{R}}
		= S^1_\beta \times [0,\tilde{L}],
	\quad
		\dd{\tilde{s}}^2
		= \frac{1}{\Omega^2} \dd{s}^2
		= \dd{\tau}^2 + \dd{\tilde{x}}^2
	\end{equation}
	This is the corresponding geometry of a thermal partition function, with a constant temperature $\beta_\tau = 2\pi$. 
	Now $\pdd{\tau}$ is truly a Killing vector of $\tilde{\mcal{R}}$. 
	In this case the thermal entropy is extensive: the total entropy $S_\mcal{A}$ is proportional to the size of the system:
	\begin{equation}
		S_\mcal{A}
%		= s\,\norm{f(\mcal{A})}_{\tilde{\mcal{R}}}
		= \frac{c}{6}\,\norm{f(\mcal{A})}_{\tilde{\mcal{R}}}
	\end{equation}
	Here $\norm{f(\mcal{A})}_{\tilde{\mcal{R}}}$ denotes the $\tilde{R}$ length of the interval $f(\mcal{A})$:
	\begin{equation}
		\norm{f(\mcal{A})}_{\tilde{\mcal{R}}}
		= \int_{f(\mcal{A})} \dd{\tilde{s}}
		= \int_{f(\mcal{A})} \frac{\dd{s}}{\Omega}
		= \int_{\mcal{A}} \frac{\dd{s}}{\Omega}
		= \int_{\mcal{A}} \frac{2\pi}{\beta(x)} \dd{x}
	\end{equation}
	We've used the fact the $f$ is a diffeomorphism, to reduce the integral back to an integral along $\mcal{A}$ in the original coordinates. 
	The entropy is thus given by:
	\begin{equation}
		S_\mcal{A}
		= \frac{c}{6}
			\int_{\mcal{A}} \frac{2\pi}{\beta(x)} \dd{x}
	\end{equation}

\pagebreak
\textit{
	The discussion above is in the framework of $\mrm{CFT}_2$, but with different modular flow generator. Now we consider the framework of $T\bar{T}$ deformed theories. The thermal entropy with unit length is
	\begin{equation}
	S_{thermal}=\frac{\pi c}{3\beta}\eta
	\end{equation}
	\begin{equation}
	\eta=\frac{1}{\sqrt{1+\frac{4\pi^2c\mu}{3\beta^2}}}
	\end{equation}
	Consider the half line case,
	\begin{equation}
	\beta(x)=2\pi x
	\end{equation}
	One can get the C-function,
	\begin{equation}
	C(l)=\frac{c}{6}\frac{1}{\sqrt{1+\frac{c\mu}{3l^2}}}
	\end{equation}
	Again, there is ``one endpoint'' in this case, which will give one half contribution of the C-function with the pre-factor $c/3$.
}
